Analysis-ready VCF at Biobank scale using Zarr

### 2025-01-22 paper.tex
Pass through the S3 section

While these benchmarks are comparable with results obtained on the 
shared file-system in the Genomics England research environment,
the sequential single-threaded access patterns used do not map
well to modern cloud computing environments.
As each request to a cloud object store has a high latency, it 
is important to have many requests running concurrently to achieve 
good performance~\citep{durner2023exploiting}.
To explore the limits of performance we developed a simple prototype
Zarr implementation, designed to maximise concurrency using asynchrous
IO and multiple processes (see Methods for details).
Fig~\ref{fig-s3-throughput} shows the rate at which genotypes 
can be decoded to RAM as well as rate at which we can compute
af-dist (using Numba to classify genotypes, as before)
as we increase the number of parallel processes on
an EC2 c5.24xlarge instance with 96 vCPUs (48 physical).
The maximum decompression rate of 47.4GiB/s 
occurs at 64 processes, with a corresponding 
network data rate of 127 MB/s [FIXME confirm and convert to MiB/s].
Fig~\ref{fig-s3-network-throughput} explores the limitations imposed by
network access by measuring the rate at which we can download the chunks
into RAM (without decompressing) using the same parallel and asynchronous
architecture. Here, the maximum throughput of

To assess the limitations imposed by network access, we repeated this
experiment but measured only the rate at which chunks can be downloaded

When calculating af-dist after
decompression the throughput was 24.6 GB/s. With this parallel approach the 
af-dist calculation above was then completed in 4.12s elapsed time.
This level of throughput would complete af-dist for all chromosomes for all samples
% compute: $4.08 * 1.27
% io: (((722000000/10000)*(79000/1000))/1000)*$0.0004
in 1.27 hours for a compute cost of \$5.18 and an io cost of \$2.28.
(See \citep{durbin2020task,moore2021ome,gowan2022using} for further Zarr
benchmarks in cloud settings.)

>>>>

While these benchmarks are comparable with results obtained on the 
shared file-system in the Genomics England research environment,
the sequential single-threaded access patterns used do not map
well to modern cloud computing environments.
As each request to a cloud object store has a high latency, it 
is important to have many requests running concurrently to achieve 
good performance~\citep{durner2023exploiting}.
To explore the limits of performance we developed a simple prototype
Zarr implementation, designed to maximise concurrency using asynchrous
IO and multiple processes (see Methods for details).
Fig~\ref{fig-s3-throughput} shows the rate at which genotypes 
can be decoded to RAM as well as rate at which we can compute
af-dist (using Numba to classify genotypes, as before)
as we increase the number of parallel processes on
an EC2 c5.24xlarge instance with 96 vCPUs (48 physical).
The maximum decompression rate of 47.4GiB/s 
occurs at 64 processes, with a corresponding 
network data rate of 127 MB/s [FIXME confirm and convert to MiB/s].
Fig~\ref{fig-s3-network-throughput} explores the limitations imposed by
network access by measuring the rate at which we can download the chunks
into RAM (without decompressing) using the same parallel and asynchronous
architecture. Here, the maximum rate of 1.2GiB/s for a c5.9xlarge instance
and 2.4GiB/s for a c5n.9xlarge (high performance network) instance
was achieved with 32 processes. Genotype data can therefore be fetched 
from S3 around an order of magnitude more quickly than it can be decompressed 
into RAM. This has some interesting consequences
when we consider the goal of seeking the highest possible compression levels 
for genotype data and the usual cloud storage cost structuring models (see
Discussion).

Returning to Fig~\ref{fig-s3-throughput}, we can see that the af-dist
calculation is performed at a maximum rate of 25.5GiB/s using 128 
processes. At this rate, the af-dist calculation for the target region 
was completed in 4.12s [CHECKME] (compare to 5.23min for the naive single-threaded
method above).
This level of throughput would complete af-dist for all chromosomes for all samples
% compute: $4.08 * 1.27
% What are the numbers here??
% io: (((722000000/10000)*(79000/1000))/1000)*$0.0004
in 1.27 hours for a compute cost of \$5.18 and an IO cost of \$2.28.
(See \citep{durbin2020task,moore2021ome,gowan2022using} for further Zarr
benchmarks in cloud settings.)
